%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Classicthesis Typographic Thesis
% LaTeX Template
% Version 1.4 (1/1/16)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% André Miede (http://www.miede.de) with commenting modifications by:
% Vel (vel@LaTeXTemplates.com)
%
% License:
% GNU General Public License (v2)
%
% General Tips:
% 1) Make sure to edit the classicthesis-config.file
% 2) New enumeration (A., B., C., etc in small caps): \begin{aenumerate} \end{aenumerate}
% 3) For margin notes: \marginpar or \graffito{}
% 4) Do not use bold fonts in this style, it is designed around them
% 5) Use tables as in the examples
% 6) See classicthesis-preamble.sty for useful commands
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
		twoside,openright,titlepage,numbers=noenddot,headinclude,%1headlines,
	 	footinclude=true,cleardoublepage=empty,
		dottedtoc, % Make page numbers in the table of contents flushed right with dots leading to them
		BCOR=5mm,paper=a4,fontsize=11pt, % Binding correction, paper type and font size
		american, % Languages, change this to your language(s)
		]{scrreprt} 
     
     %\overfullrule=1mm
                
% Includes the file which contains all the document configurations and packages - make sure to edit this file
\input{classicthesis-config}

\addbibresource{Bibliography.bib} % The file housing your bibliography
%\addbibresource[label=ownpubs]{Self_Publications.bib} % Uncomment for optional self-publications

\DeclareMathAlphabet{\mymathbb}{U}{BOONDOX-ds}{m}{n}
\SetKwRepeat{Do}{do}{while}
\SetKw{KwBy}{by}

\newcommand{\PrimaryColor}[1]{{\color{Maroon} #1}}

\newcommand{\axisStyle}[2]{title style={at={(0.5,1.1)},anchor=north, yshift=-0.1},
    title=#1,
  x tick label style={/pgf/number format/fixed},
  ymajorgrids,
  ymin=0.0,
  ymax=1.0,
  ytick={0.0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0},  
    % nodes near coords,
    width=8.5cm,
    height=7cm,
  ylabel=#2 robustness,
  xlabel=Size,
    enlarge y limits={0.1,upper},
    enlarge x limits=0.15,    
    legend pos=south east,
  ybar=1.4pt,
    bar width=10pt,}

% \newtheorem{problem2}{Problem}
% \newtheorem{definition}{Definition}

% \newcommand{\centered}[1]{\begin{tabular}{l} #1 \end{tabular}}

\NewDocumentCommand{\rot}{O{60} O{1em} m}{\makebox[#2][l]{\rotatebox{#1}{#3}}}%

% Uncomment the following line to disable todos
\usepackage[disable]{todonotes}
% \usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage{ragged2e}
\usepackage{longtable}
\usepackage{lscape}
\usepackage{marginnote}
\usepackage{siunitx}
\usepackage{tabularx}
\usepackage[export]{adjustbox}
\usepackage{microtype} 
\usepackage{cleveref}
\usepackage{multirow}
\usepackage{soul,xcolor}
\usepackage{color,colortbl}
\definecolor{Gray}{gray}{0.9}

\usepackage[acronym]{glossaries}
\makeglossaries
\newacronym{bn}{BN}{Boolean network}
\newacronym{psbn}{PSBN}{partially specified Boolean network}

% \usepackage{subcaption}
% \usepackage{caption}
% \usepackage{graphicx}

\renewcommand{\algorithmautorefname}{Algorithm}
\newcommand{\exampleautorefname}{Example}
\newcommand{\definitionautorefname}{Definition}


\input{Chapters/my_commands}

\begin{document}

\frenchspacing % Reduces space after periods to make text more compact

\raggedbottom % Makes all pages the height of the text on that page

%\selectlanguage{american} % Select your default language - e.g. american or ngerman

%\renewcommand*{\bibname}{new name} % Uncomment to change the name of the bibliography
%\setbibpreamble{} % Uncomment to include a preamble to the bibliography - some text before the reference list starts

\pagenumbering{roman} % Roman page numbering prior to the start of the thesis content (i, ii, iii, etc)

\pagestyle{plain} % Suppress headers for the pre-content pages

%----------------------------------------------------------------------------------------
%	PRE-CONTENT THESIS PAGES
%----------------------------------------------------------------------------------------

\include{FrontBackMatter/Titlepage} % Main title page

\include{FrontBackMatter/Titleback} % Back of the title page

\cleardoublepage\include{FrontBackMatter/Dedication} % Dedication page - uncomment and create a Dedication.tex to include a dedication -> moticko? :D

\cleardoublepage\include{FrontBackMatter/Declaration} % Declaration page

%\cleardoublepage\include{FrontBackMatter/Foreword} % Uncomment and create a Foreword.tex to include a foreword

\cleardoublepage\include{FrontBackMatter/Abstract} % Abstract page

\cleardoublepage\include{FrontBackMatter/Publications} % Publications from the thesis page

% \cleardoublepage\include{FrontBackMatter/ReproducibilityStatement} % Links to experiments and benchmarks

\cleardoublepage\include{FrontBackMatter/Acknowledgments} % Acknowledgements page

\cleardoublepage

\pagestyle{scrheadings} % Show chapter titles as headings

\cleardoublepage\include{FrontBackMatter/Contents} % Contents, list of figures/tables/listings and acronyms 

\cleardoublepage

\pagenumbering{arabic} % Arabic page numbering for thesis content (1, 2, 3, etc)
%\setcounter{page}{90} % Uncomment to manually start the page counter at an arbitrary value (for example if you wish to count the pre-content pages in the page count)

\cleardoublepage % Avoids problems with pdfbookmark 

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - CHAPTERS
%----------------------------------------------------------------------------------------

\include{Chapters/Introduction}

\include{Chapters/Preliminaries}

\include{Chapters/StateOfTheArt}
 
\include{Chapters/SourceTargetControl}

\include{Chapters/PhenotypeControl}

\include{Chapters/Software}

\include{Chapters/Applications}

\cleardoublepage

% Visually separate conclusions from other chapters in TOC

\chapter*{Conclusions}
\label{chapter:conclusion}
\addcontentsline{toc}{chapter}{\textsc{Conclusions}}
\addtocontents{toc}{\protect\bigskip}

In this thesis, we investigated control of partially specified Boolean
networks, a problem that had not been systematically explored before this
work. Existing control methods for Boolean networks mostly assume fully
specified update functions. This thesis closes part of this gap by
bringing control to the partially specified setting and by formulating control
objectives that explicitly account for uncertainty in the Boolean update
functions.

The first contribution of this work is formalization of source-target control
for partially specified Boolean networks. In contrast to classical Boolean
network control, where the result is typically a minimal perturbation set, the
partially specified setting requires a different view of optimality. Since a
perturbation may succeed only for some interpretations of the unknown
functions, the methods developed in this thesis compute not only candidate
perturbations, but also their robustness with respect to model uncertainty.
This provides a more informative output: instead of a single optimal
intervention, the user obtains a set of possible interventions together with
the subset of interpretations for which they are guaranteed to
work.

The second contribution is a suite of algorithms solving this problem.
The initial semi-symbolic method addressed one-step source-target control,
while the latter fully symbolic method generalized the approach to one-step,
temporary, and permanent perturbations. These methods improve the state
of the art by avoiding the explicit enumeration of all Boolean network
instances, whose number can be very large in partially specified models.
Instead, they exploit symbolic state-space representations to reason about
states, interpretations, and perturbations together. The experiments presented in this thesis
show that temporary and permanent perturbations can provide smaller and
more robust control strategies than one-step interventions, demonstrating
that the richer perturbation models are not only theoretically more general
but also practically useful.

While developing source-target control, we identified an important limitation
of attractor-based objectives in the partially specified setting. Different
interpretations of the same partially specified model may exhibit shifted or
slightly different attractors. As a consequence, requiring convergence to one
predefined target attractor can be too restrictive for biological applications,
where the exact attractor may be less important than the biological traits it
represents. To address this limitation, we introduced phenotype control, in
which the target is specified by a set of marker conditions rather than by an
explicit attractor. This formulation improves the usability of control in
biological case studies, because it allows the user to specify the desired
phenotype through characteristic traits, such as the activation or inhibition
of selected genes, without having to know the precise target attractor in
advance.

The third contribution is the implementation of the developed algorithms in
the AEON ecosystem. This part involves the symbolic manipulation of both the state space and the
space of interpretations, as well as a compact encoding of perturbations. The
implementation builds on BDD-based symbolic representations and extends
the existing AEON infrastructure with control-specific components. In
particular, perturbations are encoded in a way that allows them to be handled
within the same symbolic framework as the original partially specified model.
The methods were implemented in performant Rust libraries, exposed
through a Python interface for programmatic use, and integrated into a
web-based interface for interactive phenotype control. As a result, the thesis
does not only provide theoretical algorithms, but also makes them available
to users through practical software.

The fourth contribution is the application of the methods to biological
systems. We demonstrated the applicability of the developed control
algorithms on real-world models, including case studies related to cell
differentiation and MAPK/RTK signalling. These examples show how control
can be used not only as a tool for proposing reprogramming interventions,
but also as a tool for model refinement. In particular, control results can
identify perturbations whose experimental validation would be informative
for distinguishing between alternative interpretations of a partially specified
model. This connects the control problem back to the broader modelling
workflow: rather than treating model construction, analysis, control, and
experimental refinement as separate steps, the proposed framework uses
control to guide the refinement of the model itself.

At the same time, the work has several limitations. First, the methods focus
on variable perturbations, corresponding to fixing some variables to
constant values. Other forms of interventions, such as edge,
function, or sequential perturbations, remain outside the
scope of this thesis. Second, phenotype control is currently developed only
for permanent perturbations, while temporary phenotype control remains an
open direction. Third, although symbolic methods avoid explicit enumeration
of all interpretations, scalability is still limited by the size and structure of
the symbolic representation. Large models or models with many unknown
functions may still become computationally challenging. Fourth, the
biological interpretation of computed perturbations requires care: a
perturbation that is valid in the formal model may not always correspond to
a feasible, safe, or experimentally realizable biological intervention. Finally,
the proposed refinement workflow depends on the availability of suitable
experimental readouts that can discriminate between relevant model
interpretations.

Several directions for future research follow from these limitations. A direct
extension is phenotype control with temporary perturbations, which would
combine the biological flexibility of phenotype-based targets with less
intrusive intervention strategies. Another promising direction is the
combination of control-guided refinement with model checking, so that
control objectives and temporal specifications can be used together to select
the most informative experiments. Future work could also investigate richer
classes of perturbations, including edge- or function-level interventions, and
their relationship to biological mechanisms such as mutations, pathway
inhibition, or context-dependent regulation.

Overall, this thesis lays the foundation for control of partially specified Boolean networks
as a formal and computational framework for reasoning about interventions
under model uncertainty. By combining new problem formulations, symbolic
algorithms, practical software support, and biological case studies, it provides
a basis for using control not only to guide system behavior, but also to support
the iterative refinement of uncertain biological models.


\cleardoublepage

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - APPENDICES
%----------------------------------------------------------------------------------------

\appendix


\include{Chapters/AppendixNotation}

\include{Chapters/AppendixDigitals}

%----------------------------------------------------------------------------------------
%	POST-CONTENT THESIS PAGES
%----------------------------------------------------------------------------------------

\cleardoublepage\include{FrontBackMatter/Bibliography} % Bibliography

%\cleardoublepage\include{FrontBackMatter/Declaration} % Declaration

%\cleardoublepage\include{FrontBackMatter/Colophon} % Colophon

%----------------------------------------------------------------------------------------

\end{document}
