% Acknowledgements

\pdfbookmark[1]{Acknowledgements}{Acknowledgements} % Bookmark name visible in a PDF viewer

\begingroup

\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{\textsc{Acknowledgements}}

As no man is an island, no thesis is completed alone. I would like to express my sincere gratitude to my supervisor, Luboš Brim, for giving me the opportunity to work on this topic and for his continuous support and guidance throughout my PhD. I am equally grateful to David Šafránek for taking up the baton and for his help and support during my studies.

I would also like to thank all members of the SyBiLa group for their collaboration and for keeping me inspired. Special thanks go to Samuel Pastva and Matej Troják for sharing the hardships of doctoral studies with me and for their mutual support, to Vojtěch Veselý for development of the AEON tool user interface.

My deepest thanks go to my significant other, Michal, who has been my inspiration trough out of the whole journey. Enrolling in the PhD program at the same time and reaching the finish line sooner than I did, he has been a constant source of motivation. I am grateful for his understanding and support during the most challenging times.

Finally, I would like to thank my family for their unwavering support and encouragement throughout my studies, and my friends and colleagues for their understanding, support, and for resisting the urge to ask about the progress of my thesis.

\endgroup

