\chapter{Digital attachments}

\label{appendix:digi_attach}

All of the digital attachments related to this thesis are attached in the supplementary files. They are also available online in various repositories. The list of the digital attachments is as follows:

\section*{One-step semi-symbolic source-target control} 
The one-step semi-symbolic source-target control from \autoref{sec:target_semisymbolic} is stored in the \PrimaryColor{\texttt{src\_semisym\_control.zip}} archive. It is a legacy implementation of \texttt{biodivine-pbn-control} library and can be also found in the history of git repository\footnote{\url{https://github.com/sybila/biodivine-pbn-control/tree/e973748}}.

\section*{Source-target and phenotype control}
The implementation of methods with symbolic state space representation for source-target (\autoref{sec:ostp_symbolic}) and phenotype control (\autoref{chapter:phenotype_control_psbn}) is stored in the \PrimaryColor{\texttt{src\_symbolic\_control.zip}} archive. It is a current implementation of \texttt{biodivine-pbn-control} library and can be also found in the git repository\footnote{\url{https://github.com/sybila/biodivine-pbn-control}}.

\section*{Myeloid case study}
The case study for control of the myeloid differentiation model from \autoref{sec:myeloid_case_study} is stored in the \PrimaryColor{\texttt{myeloid\_case\_study.zip}} archive. It contains a Jupyter notebook with the code and data for the case study. It can be also found in the git repository\footnote{\url{https://github.com/sybila/biodivine-aeon-py/tree/main/example/case-study/control}}.

\section*{MAPK case study}
The case study for control of the MAPK model from \autoref{sec:mapk_case_study} is stored in the \PrimaryColor{\texttt{mapk\_case\_study.zip}} archive. It contains four Jupyter notebooks with the code and data for the case study and also with a tutorial on how to use AEON.py. It can be also found in the Zenodo archive\footnote{\url{https://doi.org/10.5281/zenodo.16886813}}.

\section*{Others}
The implementation of AEON.py described in \autoref{sec:python_package} can be found in the git repository\footnote{\url{https://github.com/sybila/biodivine-aeon-py}}. 

The implementation of the frontend of AEON described in \autoref{sec:aeon_web_app} can be found in the git repository\footnote{\url{https://github.com/sybila/biodivine-aeon-react}}.